Inequality in socioemotional skills: a cross-cohort comparison

### 2019-02-27 codes/R/paper/seskills_jpube_rev.Rnw
new version of main Rnw file

\documentclass[pdftex,10pt,a4paper]{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{amstext}% for resizing the symbol in math
\usepackage{amsthm}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage[babel]{csquotes}
\usepackage{booktabs}
\usepackage{tocloft}
\usepackage{dsfont}
\usepackage{xcolor}
\usepackage{accents}
\usepackage{rotating}
\usepackage{pdflscape}  % for landscape page
\usepackage{lmodern}  %font
\usepackage{natbib}
\usepackage{array}
\usepackage{graphicx}
\usepackage{adjustbox}
\usepackage{tikz}
\usepackage{bm}
\usepackage{array}
\usepackage{multirow}
\usepackage[multiple]{footmisc}   % for multiple footnotes
\usepackage{textgreek}            % for greek letters in text
\usepackage{ragged2e}             % for table notes justification
\usepackage{setspace} % for spacing
\onehalfspacing

@@

\usepackage{authblk}
\title{Inequality in socioemotional skills: a cross-cohort comparison}
\author{Orazio Attanasio \hspace{10pt} Richard Blundell \\  Gabriella Conti \hspace{10pt} Giacomo Mason}
\affil{\emph{Institute for Fiscal Studies} \\ \emph{Department of Economics, University College London}}
\date{October 2018}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\begin{document}

@@

\thanks{
\footnotesize{Corresponding Author: Orazio Attanasio, Department of Economics, University College London,
Gordon Street, London WC1H0AX, UK. Email: o.attanasio@ucl.ac.uk. We thank participants to the NBER/LSE Trans-Atlantic Public Economics Seminar (TAPES) and our two discussants Hilary Hoynes and Gabriel Ulyssea for excellent comments. We are grateful to The Centre for Longitudinal Studies, UCL Institute of Education, for the collection of the data, and to the UK Data Archive and UK Data Service for making them available. However, they bear no responsibility for the analysis or interpretation of the data. We're especially thankful to George Ploubidis for his guidance in the early stages of the project. Giacomo Mason is funded by an Economics and Social Research Council scholarship.}}
\end{titlepage}

@@

<<READ, cache=FALSE >>=
library(here)
read_chunk(here("codes/R/1_load.R"))
read_chunk(here("codes/R/2_efa.R"))
read_chunk(here("codes/R/3_inv_main.R"))
read_chunk(here("codes/R/3_inv_aux.R"))
read_chunk(here("codes/R/3_inv_fit.R"))
read_chunk(here("codes/R/4_bootstrap.R"))
read_chunk(here("codes/R/5_selectvars.R"))
read_chunk(here("codes/R/5_sumstats.R"))
read_chunk(here("codes/R/6_graphs.R"))
read_chunk(here("codes/R/6_graphs_ineq.R"))
read_chunk(here("codes/R/6_graphs_item.R"))
read_chunk(here("codes/R/8_regs_dep.R"))
read_chunk(here("codes/R/8_regs_outc.R"))
read_chunk(here("codes/R/99_auxfun.R"))

@@

@@

@@

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\clearpage

\begin{table}[ht]
\caption{Subscale of comparable items}\label{tab:scalecomp}
\centering\setstretch{1.2}
\begingroup\footnotesize
\begin{tabular}{C{.03\textwidth}C{.05\textwidth}C{.04\textwidth}L{.12\textwidth}L{.3\textwidth}L{.3\textwidth}}
  \toprule
 Itm. & Factor & Cat. & Title & Rutter Wording (BCS 1970) & SDQ Wording (MCS 2000/1) \\ \midrule
 1 & EXT & 3 & \textit{Restless} & Very restless. Often running about or jumping up and down. Hardly ever still & Restless, overactive, cannot stay still for long   \\
   2 & EXT & 3 & \textit{Squirmy/fidgety} & Is squirmy or fidgety & Constantly fidgeting or squirming  \\
   3 & EXT & 3 & \textit{Fights/bullies} & Frequently fights other children + \newline Bullies other children & Often fights with other children or bullies them  \\
   4 & EXT & 3 & \textit{Distracted} & Cannot settle to anything for more than a few moments & Easily distracted, concentration wanders  \\
   5 & EXT & 2 & \textit{Tantrums} & Has temper tantrums & Often has temper tantrums or hot tempers \\
   6 & EXT & 2 & \textit{Disobedient} & Is often disobedient & (+) Generally obedient, usually does what adults request \\
   7 & INT & 3 & \textit{Worried} & Often worried, worries about many things  & Many worries, often seems worried \\
   8 & INT & 3 & \textit{Fearful} & Tends to be fearful or afraid of new things or new situations & Nervous or clingy in new situations, easily loses confidence  \\
   9 & INT & 3 & \textit{Solitary} & Tends to do things on his/her own, rather solitary & Rather solitary, tends to play alone  \\
  10 & INT & 3 & \textit{Unhappy} & Often appears miserable, unhappy, tearful or distressed & Often unhappy, down-hearted or tearful \\
  11 & INT & 2 & \textit{Aches} & Complains of headaches + \newline Complains of stomach-ache or has vomited & Often complains of head- aches, stomach-ache or sickness \\
   \bottomrule
\end{tabular}
\endgroup
\justify {\scriptsize \emph{Notes}: \emph{Itm.} is item number. \emph{Factor} is the latent construct to which the item loads -- EXT is Externalising skills, INT is Internalising skills. \emph{Cat.} is the number of categories in which the item is coded -- 2 denotes a binary item (applies/does not apply) and 3 denotes a 3-category item. \emph{Title} is a short label for the item. \emph{Wording} columns show the actual wording in the scales used in each of the cohort studies. Items denoted by (+) are positively coded in the original scale.
\par}
\end{table}

\begin{table}
\caption{Quantile differences in scores}\label{tab:iqdiff}
\centering\setstretch{1.2}\footnotesize
<<INEQ_STATS, dependson="SELECTDATA", results='asis'>>=

# income 
tb <- data.frame(
  aggregate(finaldata$faminc10_infl, by=list(finaldata$cohort), FUN = function(x) quantile(x, c(.1, .25, .75, .9), na.rm=T))
)
tb$d9010 <- tb$x[,4]/tb$x[,1]
tb$d7525 <- tb$x[,3]/tb$x[,2]

tb.ext <- data.frame(
  aggregate(finaldata$EXT, by=list(finaldata$cohortsex), FUN = function(x) quantile(x, c(.1, .25, .5, .75, .9), na.rm=T))
)
tb.ext$d9010 <- abs(tb.ext$x[,"90%"]-tb.ext$x[,"10%"])
tb.ext$d7525 <- abs(tb.ext$x[,"75%"]-tb.ext$x[,"25%"])
tb.ext$d9050 <- abs(tb.ext$x[,"90%"]-tb.ext$x[,"50%"])
tb.ext$d5010 <- abs(tb.ext$x[,"50%"]-tb.ext$x[,"10%"])
rownames(tb.ext) <- tb.ext$Group.1
tb.ext <- t(tb.ext[,-c(1,2)]) # remove aggregated quantiles and transpose
tb.ext <- tb.ext[,c("BCS.M","MCS.M","BCS.F","MCS.F")]

tb.int <- data.frame(
  aggregate(finaldata$INT, by=list(finaldata$cohortsex), FUN = function(x) quantile(x, c(.1, .25, .5, .75, .9), na.rm=T))
)
tb.int$d9010 <- abs(tb.int$x[,"90%"]-tb.int$x[,"10%"])
tb.int$d7525 <- abs(tb.int$x[,"75%"]-tb.int$x[,"25%"])
tb.int$d9050 <- abs(tb.int$x[,"90%"]-tb.int$x[,"50%"])
tb.int$d5010 <- abs(tb.int$x[,"50%"]-tb.int$x[,"10%"])
rownames(tb.int) <- tb.int$Group.1
tb.int <- t(tb.int[,-c(1,2)]) # remove aggregated quantiles and transpose
tb.int <- tb.int[,c("BCS.M","MCS.M","BCS.F","MCS.F")]

# assemble and print
table <- rbind(tb.ext,tb.int)
table <- cbind(data.frame(as.matrix(rep(c("90-10", "75-25", "90-50", "50-10"),2))),
               table)

print(xtable(table,
                    align="L{0cm}L{.15\\textwidth}C{.1\\textwidth}C{.1\\textwidth}C{.1\\textwidth}C{.1\\textwidth}",
                    digits = 2
                    ),
             size="\\footnotesize",
             include.rownames=F,
             include.colnames=F,
             booktabs = T,
             add.to.row = list(pos = c(list(0,4)),
                              command = c(
                              # header
            " & \\multicolumn{2}{c}{\\textbf{Males}} & \\multicolumn{2}{c}{\\textbf{Females}} \\\\ \\cmidrule(lr){2-3}  \\cmidrule(lr){4-5} 
                       & BCS (1970) & MCS (2000/1) & BCS (1970) & MCS (2000/1)  \\\\
                      \\textbf{Externalising} \\\\ ",
                      "\\textbf{Internalising} \\\\ \\midrule "
                                    )),
             sanitize.text.function = function(x){x},
             sanitize.rownames.function = function(x){""}, # to not print row numbers
             tabular.environment = "tabular", width = "\\textwidth",
             floating = FALSE
)

@
\justify {\scriptsize \emph{Notes}: The table shows differences between quantiles of the distribution of socioemotional skills, by gender and cohort. The distribution is a factor score obtained from the factor model in \autoref{sec:methods}. 
\par}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{landscape}
\begin{table}
\caption{Determinants of Socioemotional Skills}\label{tab:determ}
\centering\footnotesize
<<REGS_ALL, results='asis', dependson="REGS_DEP">>=

@@

\clearpage
\begin{table}[]
\caption{Summary statistics}\label{tab:sumstats}
\centering\setstretch{1.2} \footnotesize
<<SUMSTATS_TAB, dependson="SUMSTATS", results='asis'>>=
require(xtable)
table <- summtab 
table <- table[,!names(table) %in% c("bcs.text", "bcs.num", "mcs.text", "mcs.num")] # drop if unneeded

  print(xtable(table,
               align="L{0cm}L{.3\\textwidth}C{.1\\textwidth}C{.1\\textwidth}C{.1\\textwidth}C{.1\\textwidth}",
               digits = 1
  ), 
  include.rownames=F,
  include.colnames=F,
  booktabs = T,
  add.to.row = list(pos = c(list(0)),
                    command = c(
                      # header
                      paste(" & 
                            \\multicolumn{2}{c}{\\textbf{Males}} & \\multicolumn{2}{c}{\\textbf{Females}}
                            \\\\ \\cmidrule(lr){2-3} \\cmidrule(lr){4-5} 
                            & BCS & MCS & BCS & MCS \\\\
                            "
                            , sep="")
                    )),
  sanitize.text.function = function(x){x},
  sanitize.rownames.function = function(x){""}, # to not print row numbers
  tabular.environment = "tabular",
  floating = FALSE
  )

@
\justify {\scriptsize \emph{Notes}: The table shows mean (standard deviation) of harmonised variables used in the analysis.
\par}
\end{table}

@@

\begin{table}
\caption{Suggested number of factors to retain}\label{tab:numfac}
\centering\setstretch{1.2}\footnotesize
<<NUMFACTAB, dependson='MCSEFARES', results='asis'>>=
# combine results
efatab <- data.frame(cbind(efatab.bcs,efatab.mcs))

@@

@
\justify {\scriptsize \emph{Notes}: The table compares the optimal number of factors suggested by different aprroaches for our novel scale; scree test based approaches (optimal coordinates, acceleration factor -- \citealp{Raiche2013}), parallel analysis \citep{Horn1965}, Kaiser's eigenvalue rule \citep{Kaiser1960}, Very Simple Structure (VSS, \citealp{Revelle1979}), Velicer Minimum Average Partial test (MAP, \citealp{Velicer1976}).
\par}
\end{table}

\begin{table}
\caption{Loadings from exploratory factor analysis}\label{tab:load}
\centering\setstretch{1.2}\footnotesize
<<EFALOADTAB, dependson='MCSEFARES', results='asis'>>=
# combine results
library(kableExtra)
loadtab <- cbind(meantab$num, meantab$title, data.frame(cbind(bcsfa$loadings, mcsfa$loadings)))
colnames(loadtab) <- c("item", "title", "bcs1", "bcs2", "mcs1", "mcs2")

loadtab %>%
  mutate_at(vars(matches("cs")), round, 3) %>%
  mutate(
    bcs1 = cell_spec(bcs1, "latex", bold = ifelse(abs(bcs1) > .25, TRUE, FALSE)),
    bcs2 = cell_spec(bcs2, "latex", bold = ifelse(abs(bcs2) > .25, TRUE, FALSE)),
    mcs1 = cell_spec(mcs1, "latex", bold = ifelse(abs(mcs1) > .25, TRUE, FALSE)),
    mcs2 = cell_spec(mcs2, "latex", bold = ifelse(abs(mcs2) > .25, TRUE, FALSE))
  ) %>%
  kable("latex", escape = F, booktabs = T, linesep = "", align=c('c','l','c','c','c'),
        col.names = c("Item", "Title", "Factor 1 (EXT)", "Factor 2 (INT)", "Factor 1 (EXT)", "Factor 2 (INT)")) %>%
  add_header_above(c(" " = 2,  "BCS (1970)" = 2, "MCS (2000/1)" = 2))

@@

@@

@@

# MAKE FIT TABLES
fit_main <- afitab(list(fa.c[[1]], fa.tl[[1]], fa.tli[[1]]), 
       c("Configural", 
         "Threshold + Loading Inv", 
         "Threshold, Loading, + Intercept Inv")
)

>>>>

\documentclass[pdftex,11pt,a4paper]{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{amstext}% for resizing the symbol in math
\usepackage{amsthm}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage[babel]{csquotes}
\usepackage{booktabs}
\usepackage{tocloft}
\usepackage{dsfont}
\usepackage{xcolor}
\usepackage{accents}
\usepackage{rotating}
\usepackage{pdflscape}  % for landscape page
\usepackage{lmodern}  %font
\usepackage{natbib}
\usepackage{array}
\usepackage{graphicx}
\usepackage{adjustbox}
\usepackage{tikz}
\usepackage{bm}
\usepackage{array}
\usepackage{multirow}
\usepackage[multiple]{footmisc}   % for multiple footnotes
\usepackage{textgreek}            % for greek letters in text
\usepackage{ragged2e}             % for table notes justification
\usepackage{setspace} % for spacing
\onehalfspacing

@@

\usepackage{authblk}
\title{Inequality in socioemotional skills: a cross-cohort comparison}
\author{Orazio Attanasio \hspace{10pt} Richard Blundell \\  Gabriella Conti \hspace{10pt} Giacomo Mason}
\affil{\emph{Institute for Fiscal Studies} \\ \emph{Department of Economics, University College London}}
\date{December 2018}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\begin{document}

@@

\scriptsize{Corresponding Author: Orazio Attanasio, Department of Economics, University College London,
Gordon Street, London WC1H0AX, UK. Email: o.attanasio@ucl.ac.uk. We thank participants to the NBER/LSE Trans-Atlantic Public Economics Seminar (TAPES) and our two discussants Hilary Hoynes and Gabriel Ulyssea for excellent comments. We're especially thankful to George Ploubidis for his guidance in the early stages of the project. This work is based on data from the British Cohort Study and the Millennium Cohort Study. We are grateful to The Centre for Longitudinal Studies, UCL Institute of Education, for the collection of the data, and to the UK Data Archive and UK Data Service for making them available. However, they bear no responsibility for the analysis or interpretation of the data. We would like to thank all cohort members and their families, who generously gave their time to participate in the surveys. We thank the DynaHEALTH consortium (H2020–633595 DynaHEALTH) for funding and support. GM is funded by an Economics and Social Research Council scholarship.}
\end{titlepage}

@@

<<READ, cache=FALSE >>=
library(here)
read_chunk(here("codes/R/1_load.R"))
read_chunk(here("codes/R/2_efa.R"))
read_chunk(here("codes/R/2b_efa_all.R"))
read_chunk(here("codes/R/3_inv_main.R"))
read_chunk(here("codes/R/3_inv_aux.R"))
read_chunk(here("codes/R/3_inv_no511.R"))
read_chunk(here("codes/R/3_inv_fit.R"))
read_chunk(here("codes/R/4_bootstrap.R"))
read_chunk(here("codes/R/5_selectvars.R"))
read_chunk(here("codes/R/5_sumstats.R"))
read_chunk(here("codes/R/6_graphs.R"))
read_chunk(here("codes/R/6_graphs_ineq.R"))
read_chunk(here("codes/R/6_graphs_item.R"))
read_chunk(here("codes/R/7_ineq_quantiles.R"))
read_chunk(here("codes/R/8_regs_dep.R"))
read_chunk(here("codes/R/8_regs_outc.R"))
read_chunk(here("codes/R/99_auxfun.R"))

@@

<<BCSEFA_ALL, dependson='LOAD_DATA'>>=
@

<<MCSEFA_ALL, dependson='LOAD_DATA'>>=
@

<<NUMFAC_ALL, dependson='MCSEFA'>>=
@

@@

<<FA_INV_AUX_NO511>>=
# robustness checks without items 5-11
@

@@

<<FA_SCORES_NO511>>=
# produce scores for model without items 5 and 11
@

@@

<<QDTABLE, dependson="SELECTDATA">>=
# make table of quantile differences
@

@@

\clearpage
\begin{table}[]
\caption{Summary statistics}\label{tab:sumstats}
\centering\setstretch{1.2} \footnotesize
<<SUMSTATS_TAB, dependson="SUMSTATS", results='asis'>>=
require(xtable)
table <- summtab_estimall

  print(xtable(table,
               align="L{0cm}L{.35\\textwidth}C{.12\\textwidth}C{.12\\textwidth}C{.12\\textwidth}C{.12\\textwidth}",
               digits = 1
  ), 
  include.rownames=F,
  include.colnames=F,
  booktabs = T,
  add.to.row = list(pos = c(list(0,10)),
                    command = c(
                      # header
                      paste(" & 
                            \\multicolumn{2}{c}{\\textbf{Estimation sample}} & \\multicolumn{2}{c}{\\textbf{Full sample}} 
                            \\\\ \\cmidrule(lr){2-3} \\cmidrule(lr){4-5} 
                            & \\textbf{BCS} & \\textbf{MCS} & \\textbf{BCS} & \\textbf{MCS} \\\\
                            & $N=9545$ & $N=5572$ & $N=14063$ & $N=11530$ \\\\
                            & Mean (SD) & Mean (SD) & Mean (SD) & Wt. Mean (SD) \\\\
                            "
                            , sep=""),
                      "\\textbf{Five-year survey} \\\\"
                    )),
  sanitize.text.function = function(x){x},
  sanitize.rownames.function = function(x){""}, # to not print row numbers
  tabular.environment = "tabular",
  floating = FALSE
  )

@
\justify {\scriptsize \emph{Notes}: The table shows the mean values of harmonised variables (and the standard deviation, for continuous ones). The \emph{estimation sample} is the subsample used in the analysis. The \emph{full sample} is the entire sample of non-multiparous children in both cohorts, residing in England at birth. Mean estimates for the full sample in the MCS cohort are weighted to account for survey design.
\par}
\end{table}

\clearpage
\begin{table}[ht]
\caption{Subscale of comparable items}\label{tab:scalecomp}
\centering\setstretch{1.2}
\begingroup\footnotesize
\begin{tabular}{C{.03\textwidth}C{.05\textwidth}C{.04\textwidth}L{.12\textwidth}L{.3\textwidth}L{.3\textwidth}}
  \toprule
 Itm. & Factor & Cat. & Title & Rutter Wording (BCS 1970) & SDQ Wording (MCS 2000/1) \\ \midrule
 1 & EXT & 3 & \textit{Restless} & Very restless. Often running about or jumping up and down. Hardly ever still & Restless, overactive, cannot stay still for long   \\
   2 & EXT & 3 & \textit{Squirmy/fidgety} & Is squirmy or fidgety & Constantly fidgeting or squirming  \\
   3 & EXT & 3 & \textit{Fights/bullies} & Frequently fights other children + \newline Bullies other children & Often fights with other children or bullies them  \\
   4 & EXT & 3 & \textit{Distracted} & Cannot settle to anything for more than a few moments & Easily distracted, concentration wanders  \\
   5 & EXT & 2 & \textit{Tantrums} & Has temper tantrums & Often has temper tantrums or hot tempers \\
   6 & EXT & 2 & \textit{Disobedient} & Is often disobedient & (+) Generally obedient, usually does what adults request \\
   7 & INT & 3 & \textit{Worried} & Often worried, worries about many things  & Many worries, often seems worried \\
   8 & INT & 3 & \textit{Fearful} & Tends to be fearful or afraid of new things or new situations & Nervous or clingy in new situations, easily loses confidence  \\
   9 & INT & 3 & \textit{Solitary} & Tends to do things on his/her own, rather solitary & Rather solitary, tends to play alone  \\
  10 & INT & 3 & \textit{Unhappy} & Often appears miserable, unhappy, tearful or distressed & Often unhappy, down-hearted or tearful \\
  11 & INT & 2 & \textit{Aches} & Complains of headaches + \newline Complains of stomach-ache or has vomited & Often complains of head- aches, stomach-ache or sickness \\
   \bottomrule
\end{tabular}
\endgroup
\justify {\scriptsize \emph{Notes}: \emph{Itm.} is item number. \emph{Factor} is the latent construct to which the item loads -- EXT is Externalising skills, INT is Internalising skills. \emph{Cat.} is the number of categories in which the item is coded -- 2 denotes a binary item (applies/does not apply) and 3 denotes a 3-category item. \emph{Title} is a short label for the item. \emph{Wording} columns show the actual wording in the scales used in each of the cohort studies. Items denoted by (+) are positively worded in the original scale.
\par}
\end{table}

\begin{table}
\caption{Quantile differences in scores}\label{tab:iqdiff}
\centering\setstretch{1.2}\footnotesize
<<INEQ_STATS, dependson="QDTABLE", results='asis'>>=

print(xtable(qdtable[-1],
                    align="L{0cm}L{.15\\textwidth}C{.15\\textwidth}C{.15\\textwidth}C{.15\\textwidth}C{.1\\textwidth}",
                    digits = 2
                    ),
             size="\\footnotesize",
             include.rownames=F,
             include.colnames=F,
             booktabs = T,
             add.to.row = list(pos = c(list(0,4)),
                              command = c(
                              # header
            " & \\multicolumn{2}{c}{\\textbf{Males}} & \\multicolumn{2}{c}{\\textbf{Females}} \\\\ \\cmidrule(lr){2-3}  \\cmidrule(lr){4-5}
                     Quantile diff. & BCS (1970) & MCS (2000/1) & BCS (1970) & MCS (2000/1)  \\\\
                      \\textbf{Externalising} \\\\ ",
                      "\\textbf{Internalising} \\\\ \\midrule "
                                    )),
             sanitize.text.function = function(x){x},
             sanitize.rownames.function = function(x){""}, # to not print row numbers
             tabular.environment = "tabular", width = "\\textwidth",
             floating = FALSE
)

@
\justify {\scriptsize \emph{Notes}: The table shows differences between quantiles of the distribution of socioemotional skills, by gender and cohort. Bootstrap confidence intervals with 1,000 repetitions are in brackets. The distribution is a factor score obtained from the factor model in \autoref{sec:methods}. These distributions are shown in \autoref{fig:dens}.
\par}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{landscape}
\begin{table}
\caption{Determinants of Socioemotional Skills}\label{tab:determ}
\centering\scriptsize
<<REGS_ALL, results='asis', dependson="REGS_DEP">>=
library(texreg)

@@

@@

\begin{table}
\caption{Suggested number of factors to retain -- 11-item scale}\label{tab:numfac}
\centering\setstretch{1.2}\footnotesize
<<NUMFACTAB, dependson='MCSEFARES', results='asis'>>=
# combine results
efatab <- data.frame(cbind(efatab.bcs,efatab.mcs))

@@

@
\justify {\scriptsize \emph{Notes}: The table compares the optimal number of factors suggested by different approaches, for our novel scale: scree test based approaches (optimal coordinates, acceleration factor -- \citealp{Raiche2013}), parallel analysis \citep{Horn1965}, Kaiser's eigenvalue rule \citep{Kaiser1960}, Very Simple Structure (VSS, \citealp{Revelle1979}), Velicer Minimum Average Partial test (MAP, \citealp{Velicer1976}).
\par}
\end{table}

\begin{table}
\caption{Loadings from exploratory factor analysis -- 11-item scale}\label{tab:load}
\centering\setstretch{1.2}\footnotesize
<<EFALOADTAB, dependson='MCSEFARES', results='asis'>>=
# combine results
library(kableExtra)
loadtab <- cbind(meantab$num, meantab$title, data.frame(cbind(bcsfa$loadings, mcsfa$loadings)))
colnames(loadtab) <- c("item", "title", "bcs1", "bcs2", "mcs1", "mcs2")

loadtab %>%
  mutate_at(vars(matches("cs")), round, 3) %>%
  mutate(
    bcs1 = cell_spec(bcs1, "latex", bold = ifelse(abs(bcs1) > .3, TRUE, FALSE)),
    bcs2 = cell_spec(bcs2, "latex", bold = ifelse(abs(bcs2) > .3, TRUE, FALSE)),
    mcs1 = cell_spec(mcs1, "latex", bold = ifelse(abs(mcs1) > .3, TRUE, FALSE)),
    mcs2 = cell_spec(mcs2, "latex", bold = ifelse(abs(mcs2) > .3, TRUE, FALSE))
  ) %>%
  kable("latex", escape = F, booktabs = T, linesep = "", align=c('c','l','c','c','c'),
        col.names = c("Item", "Title", "Factor 1 (EXT)", "Factor 2 (INT)", "Factor 1 (EXT)", "Factor 2 (INT)")) %>%
  add_header_above(c(" " = 2,  "BCS (1970)" = 2, "MCS (2000/1)" = 2))

@@

\begin{table}
\caption{Suggested number of factors to retain -- Full set of items for BCS and MCS}\label{tab:numfac_all}
\centering\setstretch{1.2}\footnotesize
<<NUMFACTAB_ALL, dependson='NUMFAC_ALL', results='asis'>>=
# combine results

efatab_all %>%
  tibble::rownames_to_column() %>%
  kable("latex", escape = F, booktabs = T, linesep = "", align=c('l','c','c'),
        col.names = c("Approach", "Rutter (BCS)", "SDQ (MCS)")) %>%
  column_spec(2:3, width = "5em")

@
\justify {\scriptsize \emph{Notes}: The table compares the optimal number of factors suggested by different approaches: scree test based approaches (optimal coordinates, acceleration factor -- \citealp{Raiche2013}), parallel analysis \citep{Horn1965}, Kaiser's eigenvalue rule \citep{Kaiser1960}, Very Simple Structure (VSS, \citealp{Revelle1979}), Velicer Minimum Average Partial test (MAP, \citealp{Velicer1976}).
\par}
\end{table}

\begin{table}
\caption{Loadings from exploratory factor analysis -- full set of BCS items}\label{tab:load_bcsall}
\centering\setstretch{1.2}\footnotesize
<<EFALOADTAB_BCSALL, dependson='BCSEFA_ALL', results='asis'>>=

bcs_loadtab_all %>%
  mutate_at(vars(starts_with("f")), function(x) round(as.numeric(as.character(x)), 3)) %>%
  mutate(
    f1 = cell_spec(f1, "latex", bold = ifelse(abs(f1) > .3, TRUE, FALSE)),
    f2 = cell_spec(f2, "latex", bold = ifelse(abs(f2) > .3, TRUE, FALSE)),
    f3 = cell_spec(f3, "latex", bold = ifelse(abs(f3) > .3, TRUE, FALSE)),
    f4 = cell_spec(f4, "latex", bold = ifelse(abs(f4) > .3, TRUE, FALSE)),
    f5 = cell_spec(f5, "latex", bold = ifelse(abs(f5) > .3, TRUE, FALSE))
  ) %>%
  kable("latex", escape = F, booktabs = T, linesep = "", align=c('c','l','c','c','c','c','c'),
        col.names = c("Item", "Title", "Factor 1", "Factor 2", "Factor 3", "Factor 4 ", "Factor 5"))

@
\justify {\scriptsize \emph{Notes}: The table displays the factor loadings obtained from exploratory factor analysis (EFA) on the full set of Rutter items from the BCS. The EFA is based on the decomposition of the polychoric correlation matrix, and uses weighted least squares and oblimin rotation. Items denoted by * are used in the 11-item scale, and items denoted by ** are combined in the 11-item scale.
\par}
\end{table}

\clearpage
\begin{table}
\caption{Loadings from exploratory factor analysis -- full set of MCS items}\label{tab:load_mcsall}
\centering\setstretch{1.2}\footnotesize
<<EFALOADTAB_MCSALL, dependson='MCSEFA_ALL', results='asis'>>=

mcs_loadtab_all %>%
  mutate_at(vars(starts_with("f")), function(x) round(as.numeric(as.character(x)), 3)) %>%
  mutate(
    f1 = cell_spec(f1, "latex", bold = ifelse(abs(f1) > .3, TRUE, FALSE)),
    f2 = cell_spec(f2, "latex", bold = ifelse(abs(f2) > .3, TRUE, FALSE)),
    f3 = cell_spec(f3, "latex", bold = ifelse(abs(f3) > .3, TRUE, FALSE)),
    f4 = cell_spec(f4, "latex", bold = ifelse(abs(f4) > .3, TRUE, FALSE)),
    f5 = cell_spec(f5, "latex", bold = ifelse(abs(f5) > .3, TRUE, FALSE))
  ) %>%
  kable("latex", escape = F, booktabs = T, linesep = "", align=c('c','l','c','c','c','c','c'),
        col.names = c("Item", "Title", "Factor 1", "Factor 2", "Factor 3", "Factor 4 ", "Factor 5"))

@
\justify {\scriptsize \emph{Notes}: The table displays the factor loadings obtained from exploratory factor analysis (EFA) on the full set of SDQ items from the BCS. The EFA is based on the decomposition of the polychoric correlation matrix, and uses weighted least squares and oblimin rotation. Items denoted by * are used in the 11-item scale.
\par}
\end{table}

@@

\begin{table}[ht]
\centering\setstretch{1.2} \footnotesize
\caption{Parameterisations for measurement invariance}\label{tab:invparam}
\begin{tabular}{L{.18\textwidth}L{.45\textwidth}L{.3\textwidth}}
\toprule
\textbf{Invariance level}     & \textbf{Description} & \textbf{Restrictions} \\ \midrule \\
Configural (WE{\textTheta})   
    & \bi[label=$\cdot$] \item Minimally restrictive model for identification \ei
    & For all groups: $\begin{array}{|l} \text{diag}(\bm{\Phi}) = \bm{I} \\
                      \bm{\kappa}=\bm{0} \\
                      \bm{\nu}=\bm{0} \\
                      \text{diag}(\bm{\Psi})=\bm{I} \end{array}$
                      \\ \\
Threshold invariance
    & \bi[label=$\cdot$] \item Restricts thresholds $\tau$ to be equal across groups
                  \item Statistically equivalent to configural (when measures have 3 categories or less)\ei
    & $\tau_{1,ci} = \tau_{1,c^\prime i}$ for all items, $\forall c, c^\prime$
                      \newline $\tau_{2,ci} = \tau_{2,c^\prime i}$ for non-binary items, $\forall c, c^\prime$
                      \newline For all groups: $\begin{array}{|l} \text{diag}(\bm{\Phi}) = \bm{I} \\ 
                                                \bm{\kappa}=\bm{0} \end{array}$
                      \newline For ref. group $A$: $\;\begin{array}{|l} \bm{\nu}_A=\bm{0} \\
                                                  \text{diag}(\bm{\Sigma}_A)=\bm{I} \end{array}$
                   \\ \\
Threshold and Loading invariance
    & \bi[label=$\cdot$] \item Restricts thresholds $\tau$ and loadings $\lambda$ to be equal across groups
                  \item Allows comparison of latent factor variances \ei
    & $\tau_{1,ci} = \tau_{1,c^\prime i}$ for all items, $\forall c, c^\prime$
                      \newline $\tau_{2,ci} = \tau_{2,c^\prime i}$ for non-binary items, $\forall c, c^\prime$
                      \newline $\lambda_{ci} = \lambda_{c^\prime i}$ for all items, $\forall c, c^\prime$
                      \newline For all groups: $\bm{\kappa}=\bm{0}$
                      \newline For ref. group $A$: $\;\begin{array}{|l} \bm{\nu}_A=\bm{0} \\
                                                  \text{diag}(\bm{\Sigma}_A)=\bm{I} \\
                                                  \text{diag}(\bm{\Phi}_A) = \bm{I} \end{array}$ 
                   \\ \\
Threshold, Loading, and Intercept invariance
    & \bi[label=$\cdot$] \item Restricts thresholds $\tau$ and loadings $\lambda$ to be equal across groups
                  \item Restricts intercepts $\nu$ to zero in both groups
                  \item Allows comparison of latent factor variances \emph{and} means \ei
    & $\tau_{1,ci} = \tau_{1,c^\prime i}$ for all items, $\forall c, c^\prime$
                      \newline $\tau_{2,ci} = \tau_{2,c^\prime i}$ for non-binary items, $\forall c, c^\prime$
                      \newline $\lambda_{ci} = \lambda_{c^\prime i}$ for all items, $\forall c, c^\prime$
                      \newline For all groups: $\bm{\nu}=\bm{0}$
                      \newline For ref. group $A$: $\;\begin{array}{|l} \bm{\kappa}_A=\bm{0} \\
                                                  \text{diag}(\bm{\Sigma}_A)=\bm{I} \\
                                                  \text{diag}(\bm{\Phi}_A) = \bm{I} \end{array}$ 
\\ \\ \bottomrule
\end{tabular}
\justify {\scriptsize \emph{Notes}: Adapted from \cite{Wu2016a}.
\par}
\end{table}

@@

# MAKE FIT TABLES
fit_main <- afitab(list(fa.c[[1]], fa.tl[[1]], fa.tli[[1]]), 
       c("Configural", 
         "Threshold + Loading Inv", 
                     "Thr. + Load. + Intercept Inv")
)

fit_agesub <- afitab(list(fa.c[[2]], fa.tl[[2]], fa.tli[[2]]), 
                   c("Configural", 
                     "Threshold + Loading Inv", 
                     "Thr. + Load. + Intercept Inv")
)
fit_no511 <- afitab(list(fa.c[[6]], fa.tl[[6]], fa.tli[[6]]), 
                   c("Configural", 
                     "Threshold + Loading Inv", 
                     "Thr. + Load. + Intercept Inv")
)

table <- rbind(fit_main,fit_agesub,fit_no511)
